Stack Safety as a Security Property

### 2020-07-06 main.tex
WBCF cleanup

Our integrity and confidentiality properties concern themselves with
data, and may hold on a wide range of programs that violate expected
control flow.  But well-bracketed control flow prevents this style of
attack by guaranteeing that every jump between functions is either a
call or a return, and each return jumps to the instruction after a
corresponding call.

\leo{Also dumping in relevant wbcf discussion part. TODO: merge}
    Some forms of security rely on protected control flow. In a classic
    example, a system call that launches a missile might be gated behind code
    that checks a password. A piece of malicious code can jump into the middle
    of the sequence and execute the system call without the appropriate password,
    and still obey integrity if the malicious code never returns to its caller.
    Integrity here protects data, not privileges. But such a jump does violate
    eager stack confidentiality if the attacker jumps beyond its own function's
    instructions without performing a designated call or returning to an
    appropriate site.

    A jump to a function's own code does not violate either, even if done using
    nefarious means such as return-oriented programming (ROP). Our model would
    permit such a jump within the function, so privilege is not elevated by
    permitting ROP in this case.

    But, our policies do enforce orthogonal control flow properties that
    protect from calls that enter outside of valid entry points and from all
    invalid returns, even if they would not escalate privilege. Most notably
    {\em well-bracketed control flow}, below.

\definition{Control Separation}
A system enjoys {\em control separation} with regard to a code map
\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\) if
for any initial state \(\mach_0\) and any adjacent pair of states
\(\mach_1\) and \(\mach_2\) in the induced trace from
\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
\(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2)\), then either
\(\callmap(\mach_1(\PCname)) = n\) for some \(n\), or
\(\retmap(\mach_1)\).

\definition{Entry Integrity}
A system enjoys {\em entry integrity} with regard to a call map
\(\callmap\) and entry map \(\entmap\) if for any initial state
\(\mach_0\) and any adjacent pair of states \(\mach_c\) and
\(\mach_e\) in the induced trace from
\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
\(\callmap(\mach_c(\PCname)) = n\) then \(\entmap(\mach_e)\).

\definition{Return Integrity}
Return integrity captures the expected behavior of marked returns with
regard to their call points. We begin by defining a coinductive
relation \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_1
\MPT_2\). Given a call map \(\callmap\) and return map \(\retmap\),
\(\MPT\) is related to \(\MPT'\) if \(\MPT'\) is a suffix of \(\MPT\)
whose head is the first unmatched return in \(\MPT\). So if \(\MPT\)
starts with a return, it is related to itself:

  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
           {\(\retmap(\mach(\PCname))\)}
      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\
        \mpstate{\mach}{\pol}\)}
  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\MPT\
              \mpstate{\mach}{\pol}\MPT\)}

If \(\MPT\) starts with a non-call, non-return state, its unmatched
return must come later:

  \judgment{\(\callmap(\mach(\PCname)) = \bot
              \hspace{.5cm} \neg \retmap(\mach(\PCname))

              \mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT'\)}
           {\(\mathit{Unmatched}\ \callmap\ \retmap\
            \mpstate{\mach}{\pol}\MPT\ \MPT'\)}

And most intricately, if \(\MPT\) starts with a call, we must find the
return that matches it, which will be the unmatched return once we
step past the call. Then the first unmatched return beyond it is the
true unmatched return.

    \judgment{\(\callmap(\mach(\PCname)) = n
              \hspace{.5cm} \neg \retmap(\mach(\PCname))\)

              \(\MPT = \mpstate{\mach}{\pol} \MPT_{\mathit{call}}\)

              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{call}}\ \MPT_{\mathit{matched}}\)

              \(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\)

              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{tail}}\ \MPT_{\mathit{unmatched}}\)}
            {\(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT_{\mathit{unmatched}}\)}

A system enjoys {\em return integrity} with regard to a call map
\(\callmap\) and return map \(\retmap\) if for any initial state
\(\mach_0\) and any call state and policy state \(\mach_c\) and
\(\pol_c\) in the induced trace from
\(\mpstate{\mach_0}{\pinit(\mach_0)}\), if \(\mpstate{\mach_c}{\pol_c}
\hookrightarrow_P \MPT\) and \(\mach_r\) is the first unmatched
return, with \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT \ \MPT'\)
and \(\mach_r\) the first machine state of \(\MPT'\), then
\(\ret{\mach_c}\ \mach_r\).

>>>>

Both the integrity and confidentiality components of local state
encapsulation concern themselves with the accessibility of data, and
may hold on a wide range of programs that violate expected control
flow. For example, if {\tt g} managed to jump in the middle of the
code of {\tt h} with its call, confidentiality and integrity would
still be satisfied. 
%
However, that is still unwanted behavior: for example, a caller (or a
callee) should not be able to bypass a password check by jumping (or
returning) in the middle of a block of code - even if such an execution
would ostensibly satisfy local state encapsulation.
%
Stack safety is traditionally associated with preventing such unwanted
flows: every jump between functions should either be a call or a
return, and each return should to the instruction after its
corresponding call.

To formally capture this aspect of stack safety, we introduce three
security properties that rely on the underlying knowledge of the
program structure (code maps, function maps, return maps, and entry
maps): {\em control separation} ensures the validity of transition
sources between functions, {\em entry integrity} ensures the validity
of call targets, and {\tt return integrity} ensures the validity of
return targets.

\definition{Control Separation}

A system enjoys {\em control separation} with regard to a function map
\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
if for any initial state \((\mach_0, \pol_0)\) and any adjacent pair
of states \((\mach_1,\pol_1)\) and \((\mach_2,\pol_2)\) in the induced
trace from \(\mpstate{\mach_0}{\pol_0})\) whose current instruction
belongs to different functions according to the function map,
either $\mach_1$ is a valid call or a valid return:

\[ \codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname)) \rightarrow
\exists n. ~ \callmap(\mach_1(\PCname)) = n \vee
\retmap(\mach_1) \]

\definition{Entry Integrity}

A system enjoys {\em entry integrity} with regard to a call map
\(\callmap\) and entry map \(\entmap\) if for any initial state
\((\mach_0, \pol_0)\) and any n-transition \((\mach_c,\pol_c)\) and
\((\mach_e, \pol_e)\) in the induced trace from
\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a member of the entry map:
\[\callmap(\mach_c(\PCname)) = n \rightarrow \entmap(\mach_e)\]

\definition{Return Integrity}

\newcommand*{\Unmatched}{\mathit{Unmatched}}

Return integrity captures the expected behavior of marked returns with
regard to their call points. Given a call map $\callmap$ and a return
map $\retmap$, we begin by defining a coinductive relation
$\Unmatched$ over two $\MPOTRACE$ $\MPT$ and $\MPT'$, if $\MPT'$ is a
suffix of $\MPT$ whose head is the first unmatched return in $\MPT$.

Formally:
\begin{itemize}
\item A trace that starts with a return is related to itself:

  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
              {\(\Unmatched ~ \mpstate{\mach}{\pol} ~ \mpstate{\mach}{\pol}\)}
  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
              {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \mpstate{\mach}{\pol}\MPT\)}

\item If a trace \(\MPT\) starts with a non-call, non-return state,
  its unmatched return must come later:

  \judgment{\(\callmap(\mach(\PCname)) = \bot
              \hspace{.5cm} \neg \retmap(\mach(\PCname))
              \Unmatched ~ \MPT ~ \MPT'\)}
           {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \MPT'\)}

\item Finally, if a trace starts with the call, then we must discard the first
  unmatched return to get to the real one:

  \judgmentbr{\(\callmap(\mach(\PCname)) = n \hspace{.5cm} \neg \retmap(\mach(\PCname)) \)}
             {\(\MPT = (m,p)\MPT_{\mathit{call}}\) \quad \(\Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{matched}}\)}
             {\(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\) \quad
              \(\Unmatched ~ \MPT_{\mathit{tail}} ~ \MPT_{\mathit{unmatched}}\)}
            {\(\Unmatched ~ \MPT ~ \MPT_{\mathit{unmatched}}\)}
\end{itemize}
A system enjoys {\em return integrity} with regard to a call map
\(\callmap\) and return map \(\retmap\) if for any initial state pair
\((\mach_0, \pol_0)\) and any state pair \((\mach_c, \pol_c)\)
corresponding to a call in its induced trace, if
\(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\)
is the first unmatched return such that for some $\pol_r$
\[ \Unmatched ~ \MPT ~ (m_r,p_r)\MPT',\]
then $\mach_r$ corresponds to a return state for $\mach_c$:
\[\ret{\mach_c}\ \mach_r\]
